Rheology of Cubic Blue Phases

### 2013-07-09 trunk/papers/bprheo/bprheo.tex
Ready to submit bprheo paper
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2048 40876441-3496-0410-a613-a2fca35b2ae2

\begin{figure}[htpb]
\includegraphics[width=0.495\textwidth]{flowcurve_bp1.pdf}\\
\caption{
Flow curve $\bar{\Pi}_{xy}(\gd)$: 
Depending on the configuration in steady shear flow five different regimes 
can be distinguished: amorphous BP network with apparent yield stress (BPI-1, solid squares); 
steady flow with periodically recurring patterns (BPI-2, open squares); 
break-up of the disclination network into an amorphous state at $Er\simeq O(1)$ (BPI-3, open circles);
cholesteric helix with axis in gradient direction, also referred to
as Grandjean texture, as well as frustrated metastable Grandjean-like states that can transform into 
a regular Grandjean texture upon increasing the flow rate (BPI-4, open triangles and solid circles);
and flow-aligned nematic state (BPI-5, solid triangles). 
The error bars represent the minimum and maximum shear stress 
during one cycle in regime BPI-2.
The two dashed lines show fits to the data in regime BPI-2 with 
$\bar{\Pi}_{xy}(\gd)\propto \gd^{\;0.84}$ (left) 
and in regime BPI-4 with 
$\bar{\Pi}_{xy}(\gd)\propto \gd^{\;1.02}$ (right), respectively.}
\label{bp1-flowcurve}
\end{figure}

@@

The Grandjean configuration does not always emerge perfectly.
Fig. \ref{bp1-4-frust} shows a time sequence of a frustrated state 
that undergoes a buckling transition.
Initially the layers happen to be along a diagonal direction. 
Because the diagonal state does not have the equilibrium layer spacing the conformation 
buckles which reduces its free energy.
Due to the periodic boundary conditions and their interlocking effect on the configuration 
this state cannot transform into a perfect Grandjean texture. However, the tilt angle of the
layers with respect to the flow-vorticity ($xz$) plane decreases significantly if the 
strain rate is increased.
Interestingly, the frustration can occur for any flow rate in regime BPI-4 as shown if Fig.~\ref{bp1-flowcurve}.
 a consequence of the shearing geometry and the relative orientation of the initial state

>>>>

\begin{figure}[htpb]
\includegraphics[width=0.495\textwidth]{flowcurve_bp1.pdf}\\
\caption{
Flow curve $\bar{\Pi}_{xy}(\gd)$: 
Depending on the configuration in steady shear flow five different regimes 
can be distinguished: amorphous BP network with apparent yield stress (BPI-1, solid squares); 
steady flow with periodically recurring patterns (BPI-2, open squares); 
break-up of the disclination network into an amorphous state at $Er\simeq O(1)$ (BPI-3, open circles);
cholesteric helix with axis in gradient direction, also referred to
as Grandjean texture (BPI-4, open triangles), as well as frustrated metastable 
Grandjean textures (BPI-4, solid circles) that can transform into 
a regular Grandjean texture upon increasing the flow rate;
and flow-aligned nematic state (BPI-5, solid triangles). 
The error bars represent the minimum and maximum shear stress 
during one cycle in regime BPI-2.
The two dashed lines show fits to the data in regime BPI-2 with 
$\bar{\Pi}_{xy}(\gd)\propto \gd^{\;0.84}$ (left) 
and in regime BPI-4 with 
$\bar{\Pi}_{xy}(\gd)\propto \gd^{\;1.02}$ (right), respectively.}
\label{bp1-flowcurve}
\end{figure}

@@

The Grandjean configuration does not always emerge perfectly.
Fig. \ref{bp1-4-frust} shows a time sequence of a frustrated state 
that undergoes a buckling transition.
Initially the layers happen to be along a diagonal direction rather than
along the gradient direction. 
Because the diagonal state does not have the equilibrium layer spacing the conformation 
buckles which reduces its free energy.
Due to the periodic boundary conditions and their interlocking effect on the configuration 
this state cannot transform into a perfect Grandjean texture. However, the tilt angle of the
layers with respect to the flow-vorticity ($xz$) plane decreases significantly if the 
strain rate is increased.
This explains as well why the frustration is more frequent for low shear rates 
in regime BPI-4, as shown in Fig.~\ref{bp1-flowcurve}.
